Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?

### 2013-12-07 writeup_and_notes/male_dependent_drive.tex
committing files

Sexual reproduction is a high-stakes event that determins which chromosomes are transmitted to the next generation.  
As a consequence of this intense competition for inclusion in the next generation, alleles that  gain a transmission advantage during reproduction 
	can succeed evolutionarily, even if they have bad effects on the population, 
	the mother whose offspring they will inhabbit, 
	and even the organism in which they reside. 
This potential generates numerous conflicts  \cite{genesInConflict} including sexual conflicts between mates \cite{sexualConflict}, 
	as well as conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so.

It seems that allowing sperm to influence the outcome of meiosis would generate a confluence of these potential conflicts -- 
	sperm could actually assist an allele that distorts female meiosis.
However, this is not the case.
We find that an allele which when in sperm help distort female meiosis in its favor, %YB [too awkward / unclear?] 
	can rarely spread through a population if it bears any cost. 
Additionally, when this self-promoting driver can spread, it cannot be maintained as a protected polymorphism, and due to its positive frequency dependence 
	it spends very little time at intermediate frequency.
As such, these type of exploitation cannot drive sustained genetic conflict and so it is unlikely
	that female oogenesis and meiosis will evolve to prevent their effect.  
This likely explains why females can delay the completion of meiosis until after fertilization 
	without exploitation by collaborations by female drivers and collaborative sperm.
%Male modification of recombination rates. \yb{trying to think through this \dots{}}

Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
So long as the transmission advantage of female meiotic drive out way the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
By contrast a sperm promotor of female drive is only effective when matched with a heterozgyote female -- meaning that when rare this allele rarely enhances female drive. 
Even worse, when it does so it will find itself in a low fitness drive homozygote. 
Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do.
This is because such drive suppressing alleles reduced their chances of forming  low fitness homozygotes.
In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
	consistent with the tenuous evidence that the non-driving $ln^-$ allele is over transmitted when fertilized by $ln^+$ sperm \cite{agulnick}.

\yb{DOES THIS GO ANYWHERE? }
 We speculate that other effects of sperm on female meiosis may also be tolerated by females as a way of ensuring meiotic fairness, 
 	for example in Drosophila and other taxa, sperm accessory gland proteins (ACPs) increase the female recombination rate \cite{SOMEthING}.
Presumably this is a consequence of the resolution of double strand breaks initiated by the stressful/toxic composition of ACPs. 
However, by increasing the recombination rate sperm may decrease the effectiveness of female meiotic drivers active during MI, 
	pleiotropically enhancing female fitness in a way (while decreasing it in many others, \cite{Partridge and Hurst})

>>>>

Sexual reproduction is a high-stakes event that determines which
determines what gets transmitted to the next generation.  
As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
	can succeed evolutionarily, even if they lower the 
        fitness of the population, 
	the mother whose offspring they will inhabit, 
	and even the organism in which they reside. 
This potential generates numerous conflicts  \cite{genesInConflict} including sexual conflicts between mates \cite{sexualConflict}, 
	as well as conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so.

It seems that allowing sperm to influence the outcome of meiosis would generate a confluence of these potential conflicts -- 
	sperm could actually assist an allele that distorts female meiosis.
However, this is not the case.
We find that an allele which acts through sperm to distort female meiosis in its favor, %YB [too awkward / unclear?] 
	can rarely spread through a population if it bears any cost. 
Additionally, when this self-promoting driver can spread, it can only
rarely 
be maintained as a protected polymorphism, and due to its positive frequency dependence 
	it spends very little time at intermediate frequency.
As such, these type of exploitation cannot drive sustained genetic conflict and so it is unlikely
	that female oogenesis and meiosis will evolve to prevent their effect.  
This likely explains why females can delay the completion of meiosis until after fertilization 
	without risking exploitation by collaborations by female
        drivers and sperm alleles.
%Male modification of recombination rates. \yb{trying to think through this \dots{}}

Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
So long as the transmission advantage of female meiotic drive out way the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
By contrast a sperm promotor of female drive is only effective when matched with a heterozgyote female -- meaning that when rare this allele rarely enhances female drive. 
Even worse, when it does so it will find itself in a low fitness drive homozygote. 
Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do.
This is because such drive suppressing alleles reduced their chances of forming  low fitness homozygotes.
In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
	consistent with the tenuous evidence that the non-driving
        $ln^-$ allele is over transmitted when fertilized by $ln^+$
        sperm \cite{agulnick, Pomiankowski1993}.

\yb{DOES THIS GO ANYWHERE? }

Males can potentially influence the outcome of meiosis even before
fertilization. 
For example in Drosophila and other taxa, sperm accessory
        gland proteins (ACPs) increase the female recombination rate
        \cite{SOMEthING}, and there is some evidence that there is
        variance across male genotypes in there effect on female
        recombination rate \cite{Stevison2012}.
Presumably this is a pleiotropic consequence of the resolution of double strand breaks initiated by the stressful/toxic composition of ACPs. 
However, a pleiotropic effect of this is to increase the recombination
rate which can decrease the effectiveness of female meiotic drivers
active during MI.

	pleiotropically enhancing female fitness in a way (while
        decreasing it in many others, \cite{Partridge and Hurst}).

 We speculate that other effects of sperm on female meiosis may also be tolerated by females as a way of ensuring meiotic fairness,
